Dissecting the causal question underlying the association between cancer and ADRD

### 2021-06-06 06_presentations/melodem2021/index.Rmd
update presentation

load(here::here("03_figs", "plots.RData"))
source(here::here("02_R", "05_plots.R"))
source(here::here("02_R", "05b_plot_causes_death.R"))
hr_results <- rio::import(here::here("02_R", "hr_results.csv"))
rd_results_ever <- rio::import(here::here("02_R", "rd_results_ever.csv"))
rd_results_tv <- rio::import(here::here("02_R", "rd_results_tv.csv"))
table_results_bounds <- rio::import(here::here("02_R", "table_results_bounds.csv")) %>%
  select(1,2,rr) %>%
  filter(Proxy != "Time to cancer",
         model != "IPTW + IPCW") %>%
  pivot_wider(names_from = model, values_from = rr)

rr <- rd_results_ever %>%
  bind_rows(rd_results_tv) %>%
  select(Proxy, model, rr, conf) %>%
  pivot_wider(names_from = conf,
              values_from = rr) %>%
  mutate(contrast = "Risk Ratio") %>%
  rename(conf.high = conf.higher) %>%
  mutate(Proxy = str_replace(Proxy, "Time-varing", "Time-varying"))

results <- hr_results %>%
  mutate(contrast = "Hazard Ratio") %>%
  filter(Proxy != "Time to cancer") %>%
  mutate(Proxy = ifelse(
    Proxy == "Ever vs. Never", "Ever vs. never", "Time-varying")) %>%
  bind_rows(rr) %>%
  mutate(Proxy = str_to_sentence(Proxy))
```

@@

@@

  + For an individual still alive in year _t_, weights represent the the product of the inverse probability of surviving in each year prior to _t_, conditional on measured common causes of death and dementia.

@@

### Results

@@

results %>%
  filter(Proxy == "Time-varying") %>%
  mutate(model = fct_relevel(model, c("Unadjusted", "IPTW", "IPTW + IPCW"))) %>%
  ggplot(aes(
    y = contrast,
    x = point,
    xmin = conf.low,
    xmax = conf.high
  )) +
  geom_vline(xintercept = 1,
             color = "#E69F00",
             size = 1.2) +
  geom_pointrange(aes(shape = contrast), color = "#0072B2") +
  geom_errorbar(aes(group = contrast),
                width = 0.2,
                color = "#0072B2") +
  geom_point(
    data = temp,
    aes(x = conf.low),
    shape = 18,
    color = "#011A5E",
    size = 3
  ) +
  geom_point(
    data = temp,
    aes(x = conf.high),
    shape = 18,
    color = "#011A5E",
    size = 3
  ) +
  facet_grid(factor(model, levels = c("Unadjusted", "IPTW", "IPTW + IPCW")) ~
                 ., switch = "y") +
  scale_x_continuous(trans = "log10") +
  labs(
    x = NULL,
    y = NULL,
    group = NULL,
    shape = NULL,
    title = "Time-varying cancer",
    subtitle = "Risk ratios \U25B2, Hazard ratios \U25CF, Bounds \U25C6" ) +
  theme(
    panel.spacing = unit(0, "lines"),
    panel.background = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.text.y = element_text(
      face = "bold",
      size = rel(1.5),
      color = "#011A5E"
    ),
    axis.text.x = element_text(size = rel(1.5), color = "#011A5E"),
    axis.line.x = element_line(color = "#011A5E"),
    strip.background = element_blank(),
    plot.title = element_text(face = "bold", size = rel(2), color = "#011A5E"),
    plot.subtitle = element_text(face = "bold", size = rel(1.5), color = "#011A5E", hjust = 0.05),
    legend.position = "none")
```

>>>>

load(here::here("03_figs", "plots.RData"))
source(here::here("02_R", "05_plots.R"))
source(here::here("02_R", "05b_plot_causes_death.R"))
hr_results <- rio::import(here::here("02_R", "hr_results.csv"))
rd_results_ever <- rio::import(here::here("02_R", "rd_results_ever.csv"))
rd_results_tv <- rio::import(here::here("02_R", "rd_results_tv.csv"))
table_results_bounds <- rio::import(here::here("02_R", "table_results_bounds.csv")) %>%
  select(1,2,rr) %>%
  filter(Proxy != "Time-varing",
         model != "IPTW + IPCW") %>%
  pivot_wider(names_from = model, values_from = rr)

rr <- rd_results_ever %>%
  bind_rows(rd_results_tv) %>%
  select(Proxy, model, rr, conf) %>%
  pivot_wider(names_from = conf,
              values_from = rr) %>%
  mutate(contrast = "Risk Ratio") %>%
  rename(conf.high = conf.higher)

results <- hr_results %>%
  mutate(contrast = "Hazard Ratio") %>%
  bind_rows(rr) %>%
  mutate(Proxy = str_replace(Proxy, "Time-varying", "Time to cancer"),
         Proxy = str_replace(Proxy, "never", "Never")) %>%
  mutate(Proxy = str_to_sentence(Proxy))
```

@@

class: middle

### Acknowledgments

- Sonja A. Swanson

- Kimberly van der Willik

- M. Arfan Ikram

@@

  + For an individual still alive in year _t_, weights represent the product of the inverse probability of surviving in each year prior to _t_, conditional on measured common causes of death and dementia.

@@

### Preliminary results

@@

results %>%
  filter(Proxy == "Time to cancer") %>%
  mutate(model = fct_relevel(model, c("Unadjusted", "IPTW", "IPTW + IPCW"))) %>%
  ggplot(aes(
    y = contrast,
    x = point,
    xmin = conf.low,
    xmax = conf.high
  )) +
  geom_vline(xintercept = 1,
             color = "#E69F00",
             size = 1.2) +
  geom_pointrange(aes(shape = contrast), color = "#0072B2") +
  geom_errorbar(aes(group = contrast),
                width = 0.2,
                color = "#0072B2") +
  geom_point(
    data = temp,
    aes(x = conf.low),
    shape = 18,
    color = "#011A5E",
    size = 3
  ) +
  geom_point(
    data = temp,
    aes(x = conf.high),
    shape = 18,
    color = "#011A5E",
    size = 3
  ) +
  facet_grid(factor(model, levels = c("Unadjusted", "IPTW", "IPTW + IPCW")) ~
                 ., switch = "y") +
  scale_x_continuous(trans = "log10") +
  labs(
    x = NULL,
    y = NULL,
    group = NULL,
    shape = NULL,
    title = "Time-varying cancer",
    subtitle = "Risk ratios \U25B2, Hazard ratios \U25CF, Bounds \U25C6" ) +
  theme(
    panel.spacing = unit(0, "lines"),
    panel.background = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.text.y = element_text(
      face = "bold",
      size = rel(1.5),
      color = "#011A5E"
    ),
    axis.text.x = element_text(size = rel(1.5), color = "#011A5E"),
    axis.line.x = element_line(color = "#011A5E"),
    strip.background = element_blank(),
    plot.title = element_text(face = "bold", size = rel(2), color = "#011A5E"),
    plot.subtitle = element_text(face = "bold", size = rel(1.5), color = "#011A5E", hjust = 0.05),
    legend.position = "none")
```
